# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r echo = TRUE}
library(lubridate)
unzip("activity.zip")

data <- read.csv("activity.csv", header=TRUE, na.strings="NA")
data$date<-ymd(data$date)


```


## What is mean total number of steps taken per day?
```{r echo = TRUE}
library(dplyr)
library(ggplot2)

data<-group_by(data, date)
total_steps <- summarize(data, total = sum(steps, na.rm=TRUE))

ggplot(data = total_steps, aes(x = total)) + 
  geom_histogram(bins = 10) +
  labs(x = "Total Number of Steps per Day")

mean <- round(mean(total_steps$total), 2)
median <- median(total_steps$total)
```

The mean and median total number of steps taken per day are: mean = `r mean`, median = `r median`

## What is the average daily activity pattern?
```{r echo = TRUE}
data<-ungroup(data)
data<-group_by(data, interval)
avg_steps <- summarize(data, avg = mean(steps, na.rm = TRUE))
avg_steps <- arrange(avg_steps, interval)
ggplot(data = avg_steps, aes(x = interval, y = avg)) + geom_line() +
  labs(x = "5-min intervals", y = "Average Daily Steps")

max_steps <- max(avg_steps$avg)
max_interval <- avg_steps[avg_steps$avg == max_steps, "interval"]
```

The 5-minute interval, on average across all the days, that contains the maximum number of steps is: `r max_interval`

## Imputing missing values
```{r echo = TRUE}
missing <- nrow(data[is.na(data$steps),])
```

The number of rows with missing data is `r missing`.

### Impute using mean for that 5-min interval and create new dataset
```{r echo = TRUE}
imputed_data<-data

for (i in avg_steps$interval){
  imputed_data[imputed_data$interval==i & is.na(imputed_data$steps),"steps"] <- avg_steps[avg_steps$interval == i, "avg"]
}

```

### histogram of the total number of steps taken each day
```{r echo = TRUE}
imputed_data<-group_by(imputed_data, date)
total_steps <- summarize(imputed_data, total = sum(steps))

ggplot(data = total_steps, aes(x = total)) + 
  geom_histogram(bins = 10) +
  labs(x = "Total Number of Steps per Day")
options(scipen=999)
mean <- round(mean(total_steps$total), 2)
median <- round(median(total_steps$total), 2)
```

The mean and median total number of steps taken per day are: mean = `r mean`, median = `r median`
These numbers are different from the estimates from the first part of the assignment.


## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
imputed_data[,"weekend"] <- "weekday"
imputed_data[wday(imputed_data$date, label = TRUE) %in% c("Sun", "Sat"),"weekend"] <- "weekend"
imputed_data$weekend<-factor(imputed_data$weekend)

imputed_data<-group_by(imputed_data, weekend, interval)
avg_steps <- summarize(imputed_data, avg = mean(steps)) %>% 
  arrange(interval)


ggplot(data = avg_steps, aes(x = interval, y = avg)) + 
  facet_grid(rows = vars(weekend)) +
  geom_line() + 
  labs(y = "Number of steps")
```